Together, Individuals Make a Difference

### 2025-01-27 论文/PAPER.tex
1

	In \textbf{Task 2}, we explored the "great coach effect" by applying \textbf{Bayesian Change Point Detection (BEAST)} to analyze instances from multiple countries. The study found a significant correlation between the points of inflection in some countries' sports performance and the tenure of a "great coach," validating the importance of this effect.

@@

By analyzing the performance of various countries in different events across past Olympic Games, we extracted a feature variable $\gamma$ (\textbf{Event Adeptness Coefficient}), to represent the advantages of these countries in certain events.

@@

By extracting features, we mined the level of expertise of each country in different events within the dataset, such as China's performance across various events, as shown in the figure below:

@@

\section{Task1:Medal Prediction Model Based on LightGBM}
\subsection{Medal Table}
The medal table is influenced by various factors. To predict the standings for the next Olympics, we considered key factors identified during data preprocessing, such as nation level, participating athletes, country-specific event strengths, and adjustments to medal counts.

@@

The optimized loss function based on GBDT is represented as \textbf{squared error} in the medal prediction regression task.

@@

\begin{longtable}{l|l|l|l|l} 
	\caption{Model Performances(LightGBM)}\\ 
	\hline
	\multicolumn{1}{c|}{\textbf{Model}} & \multicolumn{1}{c|}{\textbf{MSE}} & \multicolumn{1}{c|}{\textbf{RMSE}} & \multicolumn{1}{c|}{\textbf{MAE}} & \multicolumn{1}{c}{\(\mathbf{R}^2\)}  \endfirsthead 
	\hline
	Gold Prediction                     & 0.0235                            & 0.180                              & 0.030                             & 0.890                              \\ 
	\hline
	Silver Prediction                   & 0.0239                            & 0.185                              & 0.033                             & 0.825                              \\ 
	\hline
	Bronze Prediction                   & 0.0231                            & 0.187                              & 0.034                             & 0.806                              \\
	\hline
\end{longtable}

@@

\section{Task2:Great Coach Effect}
\subsection{Verification of the Great Coach Effect}
The Great Coach effect can cause significant differences in the performance of relevant countries in corresponding events. To collect evidence of changes caused by this effect, we use a \textbf{Bayesian Changepoint Detection (BEAST)} model.

@@

Using Bayes' theorem, we can compute the \textbf{posterior probability} for each possible change point $c_k$:

@@

%5.1 排球
\begin{figure}[H]
	\centering
	\includegraphics[width=12cm]{img/Volleyball.png}
	\caption{Medal Count of Women's Volleyball Teams(CHN vs USA) Over Time}
	\label{fig:aa}
\end{figure}
Clearly, after Lang Ping began coaching the U.S. team in 2005, the U.S. volleyball team showed remarkable improvement in the next Olympic Games, while the Chinese team experienced significant regression. However, after Lang Ping returned to coach China in 2013, the Chinese team quickly rose from the trough and regained their peak performance, while the U.S. team showed a decline.

@@

%三个国家某项目衰弱 需要教练图
\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{.32\textwidth}
		\includegraphics[width=\textwidth]{img/Decline1.png}
		\caption{France}\label{subfig:1}
	\end{subfigure}
	\hfill 
	\begin{subfigure}[b]{.32\textwidth}
		\includegraphics[width=\textwidth]{img/Decline2.png}
		\caption{Sweden}\label{subfig:2}
	\end{subfigure}
	\hfill 
	\begin{subfigure}[b]{.32\textwidth}
		\includegraphics[width=\textwidth]{img/Decline3.png}
		\caption{Germany}\label{subfig:3}
	\end{subfigure}
	\caption{Three images}\label{fig:subfigures}
\end{figure}

@@

\section{Task3:Other Original Insights}

@@

LightGBM comes with a feature importance function that identifies the features with the greatest impact on the model’s predictions. By using \textbf{\textit{plot\_importance()}} function, we can visualize the importance of each feature.

@@

Furthermore, we can use Shapley value analysis to assess the model's sensitivity. \textbf{SHAP} (\textbf{SH}apley \textbf{A}dditive ex\textbf{P}lanations) is a game theory-based model explanation method used to interpret machine learning model predictions. It assigns each feature a contribution to the model's prediction, helping us understand the importance and influence of each feature in the given prediction result.The formula for Shapley value is as follows:

@@

%三个奖牌的图像
\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{.32\textwidth}
		\includegraphics[width=\textwidth]{img/shap1.png}
		\caption{Gold SHAP Summary Plot}\label{subfig:1}
	\end{subfigure}
	\hfill 
	\begin{subfigure}[b]{.32\textwidth}
		\includegraphics[width=\textwidth]{img/shap2.png}
		\caption{Silver SHAP Summary Plot}\label{subfig:2}
	\end{subfigure}
	\hfill 
	\begin{subfigure}[b]{.32\textwidth}
		\includegraphics[width=\textwidth]{img/shap3.png}
		\caption{Bronze SHAP Summary Plot}\label{subfig:3}
	\end{subfigure}
	\caption{SHAP Summary Plots}\label{fig:subfigures}
\end{figure}
From the figures, we can see that the SHAP algorithm amplifies the prior probability of winning a particular medal. In other words, athletes who have previously won gold medals have a higher probability of winning gold again if they continue to compete, and the same holds true for silver and bronze medalists. This differs somewhat from the results obtained through LightGBM's self-analysis; however, both analysis methods assign relatively high weights to Adeptness, Level, and Win Probability.

@@

4. The Bayesian Change Point Detection used in Task 2 accurately reflects the change points in time-series data. Our tests showed that the change points in the performance of China and the United States women's volleyball teams coincided with the coaching periods of Lang Ping, demonstrating the model's accuracy.
\subsection{Weaknesses}
1. Since the training data includes a large amount of historical data (spanning over 50 years), the prediction results are susceptible to the influence of historical data, leading to potential errors.

@@

We applied data preprocessing and feature engineering, using the \textbf{KMeans++} clustering algorithm to rank countries based on various features, including proficiency in different sports, host nation advantages, and athlete performance levels.

@@

\clearpage
\section{Appendix B: Program Codes}
Here are the program codes we used in our research.

@@

X.fillna(0,inplace=True)
from sklearn.preprocessing import StandardScaler
scaler=StandardScaler()
X=scaler.fit_transform(X)
columns=['host','med_d_ath','adept','lvl','win_prob','G_prob','S_prob','B_prob','lose_prob']
from sklearn.preprocessing import PolynomialFeatures
poly=PolynomialFeatures(degree=2)

>>>>

	In \textbf{Task 2}, we explored the "great coach" effect by applying \textbf{Bayesian Change Point Detection (BEAST)} to analyze instances from multiple countries. The study found a significant correlation between the points of inflection in some countries' sports performance and the tenure of a "great coach", validating the importance of this effect.

@@

By analyzing the performance of various countries in different sports across past Olympic Games, we extracted a feature variable $\gamma$ (\textbf{Event Adeptness Coefficient}), to represent the advantages of these countries in certain events.

@@

By extracting features, we mined the level of expertise of each country in different events within the dataset, such as China's performance across various sports, as shown in the figure below:

@@

\section{Task 1:Medal Prediction Model Based on LightGBM}
\subsection{Medal Table}
The medal table is influenced by various factors. To predict the standings for the next Olympics, we considered key factors identified during data preprocessing, such as nation level, participating athletes, country-specific event strengths, and adjustments to medal counts.

@@

The optimized loss function based on \textbf{GBDT}\cite{2} is represented as \textbf{squared error} in the medal prediction regression task.

@@

\begin{longtable}{l|l|l|l|l} 
	\caption{Model Performances (LightGBM)}\\ 
	\hline
	\multicolumn{1}{c|}{\textbf{Model}} & \multicolumn{1}{c|}{\textbf{MSE}} & \multicolumn{1}{c|}{\textbf{RMSE}} & \multicolumn{1}{c|}{\textbf{MAE}} & \multicolumn{1}{c}{\(\mathbf{R}^2\)}  \endfirsthead 
	\hline
	Gold Prediction                     & 0.0235                            & 0.180                              & 0.030                             & 0.890                              \\ 
	\hline
	Silver Prediction                   & 0.0239                            & 0.185                              & 0.033                             & 0.825                              \\ 
	\hline
	Bronze Prediction                   & 0.0231                            & 0.187                              & 0.034                             & 0.806                              \\
	\hline
\end{longtable}

@@

\section{Task 2:Great Coach Effect}
\subsection{Verification of the Great Coach Effect}
The Great Coach effect can cause significant differences in the performance of relevant countries in corresponding events. To collect evidence of changes caused by this effect, we use a \textbf{Bayesian Changepoint Detection (BEAST)} model.

@@

Using Bayes' theorem\cite{1}, we can compute the \textbf{posterior probability} for each possible change point $c_k$:

@@

%5.1 排球
\begin{figure}[H]
	\centering
	\includegraphics[width=12cm]{img/Volleyball.png}
	\caption{Medal Count of Women's Volleyball Teams (CHN vs USA) Over Time}
	\label{fig:aa}
\end{figure}
Clearly, after Lang Ping began coaching the U.S. team in 2005, the U.S. volleyball team showed remarkable improvement in the next Olympic Games, while the Chinese team experienced significant regression. However, after Lang Ping returned to coach China in 2013, the Chinese team quickly rose from the trough and regained their peak performance, while the U.S. team showed a decline.

@@

%三个国家某项目衰弱 需要教练图
\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{.32\textwidth}
		\includegraphics[width=\textwidth]{img/Decline1.png}
		\caption{France}\label{subfig:1}
	\end{subfigure}
	\hfill 
	\begin{subfigure}[b]{.32\textwidth}
		\includegraphics[width=\textwidth]{img/Decline2.png}
		\caption{Sweden}\label{subfig:2}
	\end{subfigure}
	\hfill 
	\begin{subfigure}[b]{.32\textwidth}
		\includegraphics[width=\textwidth]{img/Decline3.png}
		\caption{Germany}\label{subfig:3}
	\end{subfigure}
	\caption{Declining Sports in Different Countries}\label{fig:subfigures}
\end{figure}

@@

\section{Task 3:Other Original Insights}

@@

LightGBM comes with a feature importance function that identifies the features with the greatest impact on the model’s predictions\cite{2}. By using \textbf{\textit{plot\_importance()}} function, we can visualize the importance of each feature.

@@

Furthermore, we can use Shapley value analysis\cite{4} to assess the model's sensitivity. \textbf{SHAP} (\textbf{SH}apley \textbf{A}dditive ex\textbf{P}lanations) is a game theory-based model explanation method used to interpret machine learning model predictions. It assigns each feature a contribution to the model's prediction, helping us understand the importance and influence of each feature in the given prediction result. The formula for Shapley value is as follows:

@@

%三个奖牌的图像
\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{.32\textwidth}
		\includegraphics[width=\textwidth]{img/shap1.png}
		\caption{Gold SHAP Summary Plot}\label{subfig:1}
	\end{subfigure}
	\hfill 
	\begin{subfigure}[b]{.32\textwidth}
		\includegraphics[width=\textwidth]{img/shap2.png}
		\caption{Silver SHAP Summary Plot}\label{subfig:2}
	\end{subfigure}
	\hfill 
	\begin{subfigure}[b]{.32\textwidth}
		\includegraphics[width=\textwidth]{img/shap3.png}
		\caption{Bronze SHAP Summary Plot}\label{subfig:3}
	\end{subfigure}
	\caption{SHAP Summary Plots}\label{fig:subfigures}
\end{figure}
From the figures, we can see that the SHAP algorithm amplifies the prior probability of winning a particular medal\cite{4}. In other words, athletes who have previously won gold medals have a higher probability of winning gold again if they continue to compete, and the same holds true for silver and bronze medalists. This differs somewhat from the results obtained through LightGBM's self-analysis; however, both analysis methods assign relatively high weights to Adeptness, Level, and Win Probability.

@@

4. The Bayesian Change Point Detection used in Task 2 accurately reflects the change points in time-series data\cite{1}. Our tests showed that the change points in the performance of China and the United States women's volleyball teams coincided with the coaching periods of Lang Ping, demonstrating the model's accuracy.
\subsection{Weaknesses}
1. Since the training data includes a large amount of historical data (spanning over 50 years), the prediction results are susceptible to the influence of historical data, leading to potential errors.

@@

We applied data preprocessing and feature engineering, using the \textbf{KMeans++} clustering algorithm\cite{3} to rank countries based on various features, including proficiency in different sports, host nation advantages, and athlete performance levels.

@@

\clearpage
\section{Appendix B: LightGBM Training Codes}
Here are the codes we used in our LightGBM model training.

@@

X.fillna(0,inplace=True)
from sklearn.preprocessing import StandardScaler
scaler=StandardScaler()
X=scaler.fit_transform(X)
columns=['host','med_d_ath','adept','lvl','win_prob',
'G_prob','S_prob','B_prob','lose_prob']
from sklearn.preprocessing import PolynomialFeatures
poly=PolynomialFeatures(degree=2)
